---
title: "LI-5350A_DIC_Analysis"
author: "Tianyin Ouyang"
date: "2026-02-20"
output: html_document
---

This R script is written to analyze CSV data files generated by the DIC C5-12T v4.5 software using the Apollo SciTech LI-5350A DIC analyzer (LI-COR Environmental, USA). The code is specifically intended for files uploaded to Google Drive, but it can also be used with locally saved files.

##Install and Load All Required Packages 
```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("readr")) install.packages("readr")

library(tidyverse)
library(ggplot2)
library(readr)
```

##Read data files from Google Drive
A publicly accessible link to the data file must be obtained from Google Drive by clicking “Get link” and changing the access setting to “Anyone with the link can view.”
Identify the file ID from the link. The file ID is the long string located between /d/ and /view. Replace the example code with your own file ID.

For example: 
https://drive.google.com/file/d/1wINgFaJyzGOTKv5hurTC9eaPycqkR_TZ/view?usp=share_link
The file ID: "1wINgFaJyzGOTKv5hurTC9eaPycqkR_TZ"

```{r}
file_id <- "1wINgFaJyzGOTKv5hurTC9eaPycqkR_TZ"

url <- paste0("https://drive.google.com/uc?export=download&id=", file_id)
data <- read_delim(url, delim = ";",  show_col_types = FALSE)

view(data)
```

##Clean and reorganize the data table 
This step removes unnecessary columns and rows for the purpose of calculating [DIC].
```{r}
data <- data.frame(data[,c(3,5,14,17)])
data <- subset(data, data$Status != "Excluded")
view(data)
```

##Construct the calibration curve 
```{r}
std_conc <- 3000 ##replace with [DIC] (uM) in your standards

##calculate means of each calibration curve points
Cal <- aggregate(`Area..net.`~`Sample.Name` + `Volume..ml.`, data = data[1:9,], FUN = mean)

##linear regression fit of the calibration curve
fit <- lm(Cal$Area..net.~ I(std_conc*(Cal$Volume..ml./1000)))
##check information of the linear regression fit
summary(fit)

##plot the calibration curve 
eq <- paste0("y = ",round(coef(fit)[2],2), " x + ",round(coef(fit)[1],2))

ggplot(Cal,aes(x = std_conc*(Volume..ml./1000), y = Area..net.)) + geom_point(size = 3) + geom_smooth(method = "lm", linetype = "dashed", linewidth = 1.5, color = "lightblue3") + theme_bw() + theme(panel.grid = element_line(color = "gray", linewidth = 0.2, linetype = "dashed"), axis.title = element_text(size = 12, face = "bold"), axis.text = element_text(size = 12)) + ylab("Instrumental Signals") + xlab(expression(bold("[DIC] "*mu*"M"))) + annotate(geom = "text", label = eq, x = -Inf, y = Inf, hjust = -0.1, vjust = 2,fontface = "bold",size = 4.5, color = "navy")
```

