---
title: "Shimadzu_TOC_Analysis"
author: "Tianyin Ouyang"
date: "2026-04-27"
output: github_document
---
## Introduction 
This code is created to determine [NPOC] and [TDN] in water samples analyzed via Shimadzu TOC-L and TNM-L analyzer

## Install and load required packages 
```{r, messgae = FALSE}
if (!require("readr")) install.packages("readr")
if (!require("ggplot2")) install.packages("ggplot2")

library(readr)
library(ggplot2)
source("https://raw.githubusercontent.com/touyang98/General-Data-Analysis-Codes/main/Shimadzu_TOC/Shimadzu_TOC_Functions.R")
```

## Read Shimadzu Data
Export Shimadzu data to txt including all details
```{r, message = FALSE}
data <- read_delim("TOC_Test_Data.txt", delim = "\t", skip = 10)

##remove the excluded data
data <- subset(data , data$Excluded == 0)

##optional: remove unwanted or failed data 
data <- data[c(43:69,85:249),]

##categorize the data into NPOC and TDN 
data_NPOC <- subset(data, data$`Analysis(Inj.)` == "NPOC")
data_TDN <- subset(data, data$`Analysis(Inj.)` == "TN")
```

## Construct the calibration curve 
The extract_cal function is written to extract calibration curve standards from the raw data and to calculate the mean for each calibration points. 
```{r}
##NPOC calibration curve 
std_conc_NPOC <- c(5, 2.5, 1, 0.5, 0.1)
data_NPOC_cal <- extract_cal(data = data_NPOC, std_conc = std_conc_NPOC)
fit_NPOC <- lm(data_NPOC_cal$Area ~ data_NPOC_cal$std_conc)
plot_cal(data = data_NPOC_cal, fit = fit_NPOC)

##TDN calibration curve 
std_conc_TDN <- c(5, 2.5, 1, 0.5, 0.2)
data_TDN_cal <- extract_cal(data = data_TDN, std_conc = std_conc_TDN)
fit_TDN <- lm(data_TDN_cal$Area ~ data_TDN_cal$std_conc)

##optional: what if my standard injection volume is not matching sample injection volume
##for example: standard injection V is 40 uL, whereas the sample injection V is 100 uL. 
data_TDN_cal$Area <- data_TDN_cal$Area/(40/100)
```

