---
title: "Aqualog_PARAFAC_Analysis"
author: "Tianyin Ouyang"
date: "2026-02-24"
output: github_document
---

## Introduction 

## Install and load required packages 
```{r install packages, message = FALSE}
if (!require("magrittr")) install.packages("magrittr")
if (!require("staRdom")) install.packages("staRdom")
if (!require("dplyr")) install.packages("dplyr")
if (!require("writexl")) install.packages("writexl")

library(staRdom)
library(magrittr)
library(writexl)
library(dplyr)
source("https://raw.githubusercontent.com/touyang98/General-Data-Analysis-Codes/main/Aqualog_PARAFAC/Abs_Analysis_Functions.R")
source("https://raw.githubusercontent.com/touyang98/General-Data-Analysis-Codes/main/Aqualog_PARAFAC/EEMs_Analysis_Functions.R")
```

## Read absorbance data and  evaluate abs parameters
```{r read files}
## read absorbance data 
path_abs <- "PARAFAC_Data/Raw_Abs"
data_abs <- read_abs(path_abs)

## blank correction 
data_abs_corr <- blank_subtract(blank = 1, data = data_abs)

## evaluate all abs parameters 
abs_parm <- abs_parm_plus(data = data_abs_corr)
```

## Read and initial process of eems data 
```{r}
path_eems <- "PARAFAC_Data/Raw_EEMs" ##replace the example with the path of your eems data folder/ file
data_eems <- eem_read(path_eems, recursive = TRUE, import_function = "aqualog")

## blank and baseline correction
## note that the code will automatically delete the blank from the dataset
data_eems_corr <- eem_corr(data = data_eems, blank = 1)
```

## Remove scattering 
```{r}
## set what types of scattering to remove 
## each represents raman1, raman2, rayleigh1, and rayleigh2
remove_scatter <- c(TRUE, TRUE, TRUE,TRUE)

## set area of the scattering (i.e., water mark)
remove_scatter_width <- c(15,20,15,20)

## run eem_rem_scat
data_eem_rem <- eem_rem_scat(data_eems_corr, remove_scatter = remove_scatter,remove_scatter_width = remove_scatter_width)

## visualize the scattering remove 
eem_overview_plot(data_eem_rem, spp=9, contour = TRUE)
```

## Interpolate the removed scattering 
```{r}
## interpolate the gaps (calculating the missing data based on mathematics)
data_eem_interp <- eem_interp(data_eem_rem, cores = 8, type = 1, extend = FALSE)

## visualize the data
eem_overview_plot(data_eem_interp, spp=9, contour = TRUE)
```

## Smooth the data and evaluate the eems parameters
```{r, message = FALSE, warning = FALSE}
## smooth the data using eem_smooth function
data_eem_smooth <- eem_smooth(data_eem_interp)

## evaluate the eems parameter using eem_indice function written in EEMs_Analysis_Functions.R
eem_indices <- eem_indice(data = data_eem_interp, output_name = eem_indices)
```

## PARAFAC Model
```{r, message=FALSE, warning = FALSE}
## turn on the PARAFAC model in staRdom package
data(pf_models)

## input minimum and maximum component numbers 
dim_min <- 3
dim_max <- 4

## note: decrease ctol and nstart, and increase maxit can increase the model accuracy 
nstart <- 25 # number of similar models from which best is chosen
maxit = 5000 # maximum number of iterations in PARAFAC analysis 
ctol <- 10^-6 # tolerance in PARAFAC analysis

# calculating PARAFAC models
pf1 <- eem_parafac(data_eem_smooth, comps = seq(dim_min,dim_max), normalise = FALSE, const = c("uncons", "uncons", "uncons"), maxit = maxit, nstart = nstart, ctol = ctol, cores = 8)

# same model but using non-negative constraints
pf1n <- eem_parafac(data_eem_smooth, comps = seq(dim_min,dim_max), normalise = FALSE, const = c("nonneg", "nonneg", "nonneg"), maxit = maxit, nstart = nstart, ctol = ctol, cores = 8)
# rescale B and C modes to a maximum fluorescence of 1 for each component
pf1 <- lapply(pf1, eempf_rescaleBC, newscale = "Fmax")
pf1n <- lapply(pf1n, eempf_rescaleBC, newscale = "Fmax")
```

## visualize the data 
```{r, message = FALSE}
## visualize the data 
eempf_compare(pf1n, contour = TRUE)
```

## normalize by the maximum fluorescence intensity 
```{r, message = FALSE}
## change the normalization setting
pf2 <- eem_parafac(data_eem_smooth, comps = seq(dim_min,dim_max), normalise = TRUE, const = c("nonneg", "nonneg", "nonneg"), maxit = maxit, nstart = nstart, ctol = ctol, cores = 8)

## rescale B and C modes 
pf2 <- lapply(pf2, eempf_rescaleBC, newscale = "Fmax")
eempf_plot_comps(pf2, contour = TRUE, type = 1)

## visualize the data 
eempf_plot_comps(pf2, contour = TRUE, type = 1)
```

## calculate the leverage and identify outliers 
```{r}
## leverage calculation
cpl <- eempf_leverage(pf2[[1]])

## visualize the data 
eempf_leverage_plot(cpl,qlabel=0.1)
```

## Exclude outliers and rerun the PARAFAC model
```{r}
## construct a list of parameters that you would like to exclude from eems data list 
exclude <- list("ex" = c(),
                "em" = c(),
                "sample" = c("B1S2Bowers_PierPEM"))

## exclude outliers or potentially unwanted data at identified wavelengths 
eem_list_ex <- eem_exclude(eem_list, exclude)

## rerun the PARAFAC model with outlier-excluded data 
## monitor the ctol, nstart, and maxit to increase accuracy of the model 
pf3 <- eem_parafac(eem_list_ex, comps = seq(dim_min,dim_max), normalise = TRUE, maxit = maxit, nstart = nstart, ctol = ctol, cores = 8)
pf3 <- lapply(pf3, eempf_rescaleBC, newscale = "Fmax")

## visualize each components in the model 
eempf_comp_load_plot(pf3[[1]], contour = TRUE)
## note that PARAFAC model requires at least 50 data sets to model good components in EEMs; this is the reason why the test data set generate weird components
```

## split-half analysis 
```{r, eval=FALSE, echo=TRUE}
sh <- splithalf(eem_list_ex, comps = 4, normalise = TRUE, rand = FALSE, cores = 8, nstart = nstart, strictly_converging = TRUE, maxit = maxit, ctol = ctol)

splithalf_plot(sh)
```

## extract PARAFAC model output to a dataframe
```{r}
data_PARAFAC <- PARAFAC_output(pf3, data_PARAFAC, 1)
```

## export the PARAFAC model output to local computer 
```{r, eval=FALSE, echo=TRUE}
path_PARAFAC <- "Your_File_Path" ## replace the example with the path where you would like to save the file 
write_xlsx(data_PARAFAC, path_PARAFAC)
```
