---
title: "ICP-OES_Cations_Analysis"
author: "Tianyin Ouyang"
date: "2026-07-08"
output: github_document
---

## Install and Load All Required Packages 
```{r, message = FALSE, warning = FALSE}
if (!require("XML")) install.packages("XML")
if (!require("dplyr")) install.packages("dplyr")
if (!require("writexl")) install.packages("writexl")

library(dplyr)
library(XML)
library(writexl)
source("https://raw.githubusercontent.com/touyang98/General-Data-Analysis-Codes/main/iCAP_ProSeries_ICP-OES/ICP-OES_Functions.R")
```

## Read data files from local computer
```{r, message = FALSE, warning = FALSE}
path <- "Test_Data_Ouyang.xml" ##replace with your file path

data_raw <- read_xml(path = path)
```

## Extra instrumental signal data
The extract_inst function is written to clean the spreadsheet and extract instrumental signals of all analyzed elements from the raw data.
```{r}
data_ins <- extract_inst(data = data_raw)
```

## Identify elements, construct calibration curve, and evalute concentrations individually 
The extract_ion function is written to extract the interested element from the raw data. 
The cal_curve is written to construct the calibration curve of interested element. 
The sample_conc is written to calculate the concentration of interested element in samples based on the calibration curve and entered dilution fatcor. 
```{r}
## extract dataset that contain interested element individually 
ion <- "Ca"
data_Ca <- extract_ion(ion = ion, data = data_ins)

## construct the calibration curve individually 
std_row <- c(2:8) ## monitor the standard inputs for the best calibration curve
fit <- cal_curve(data = data_Ca, std_row = std_row, iFR_col = 2)## There are radical or axial iFR columns, choose one for analyses
summary(fit)

## evaluate concentrations 
dilution_factor <- 200
data_Ca$conc <- sample_conc(fit = fit, dilution_factor = dilution_factor, data = data_Ca, iFR_col = 2)

```

## Calculate concentrations of all elements 
The analyze_list function is written specifically to Pain Lab protocol, which can simutaneously calculate the concentration of 26 elements analyzed by Pain's ICP-OES instrument. 
```{r, message=FALSE, warning = FALSE}
data_analyzed <- analyze_list(data = data_ins, dilution_factor = 200)
```

## Export the data to local computer 
```{r, eval=FALSE, echo=TRUE}
path <- "Local_File_Path" ##replace with your local file path
name <- "Your_File_Name.xlsx" ##replace with whatever you would like to name your file

path_xlsx <- paste0(path, "/", name)
write_xlsx(data_analyzed, path = path_xlsx)
```
